Demes: a specification for interchanging complex demography

### 2022-02-22 paper.tex
Controversial text addition

Population sizes are given as the number of individuals, and details
such as ploidy levels are considered external to the model.
We focus on the number of individuals as opposed to the number of genomes
Sizes and mating system details are specified for each population within
epochs.
Epochs are contiguous time intervals that define
the existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or function defined by start
and end sizes that must remain positive.

@@

This model is, by design, very limited. We do not specify any details of
individuals or of the processes that happen \emph{within} populations,
because such details are necessarily complicated, vary greatly by
application, and any attempt to encompass such a broad range of biological
processes is doomed to failure. An ``individual'' (i.e., a discrete
organism) is an abstraction that should have a well defined interpretation
in most cases.
This narrow focus on individuals and their groupings into populations
should ensure that the Demes standard is stable, and not subject to
continual development as more and more elaborations of basic processes
are added. We could never hope to fully encompass all the details
of every possible simulation (including ploidy,
genome sizes, recombination, mutation and gene conversion rates,
selection coefficients, and so on) and so we clearly demarcate the
% FIXME Not quite true here, we're including selfing and cloning rate
role of the specification by including \emph{none} of them.
Such details can be specified by existing mechanisms used by
simulators. If they vary according to the demography (for example,
we have different recombination maps across populations), then
we can unambiguously define these by referring to the population
identifiers and times defined by the Demes model. Rather than attempting
to embed this information \emph{in} the demographic model, we
can define it externally and refer \emph{to} the demography.

>>>>

Population sizes are given as the number of individuals.
We focus on the number of individuals as opposed to the number of genomes
in order to not limit other software (need term -- "consumers"?) by placing potentially arbitrary
restrictions on how the numbers of chromosomes, etc., vary amongst individuals.
\jkcomment{As opposed to genomes. Why did we do this?}
\aprcomment{With this, we might want to be able to specify ploidy in as
a graph-level option, which defaults to 2. That would make the reasons
behind this decision obvious then, I think.}
\krtcomment{The number of genomes can be a non-obvious function of the number of individuals, too.
Polymorphism for chromosome number, etc., are all real complications.}
\krtcomment{We support limited aspects of the mating system--more on this below.}
Sizes and mating system details are specified for each population within
epochs. Epochs are contiguous time intervals that define

Epochs are contiguous time intervals that define
the existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or function defined by start
and end sizes that must remain positive.

@@

This model is, by design, very limited. We do not specify any details of
individuals or of the processes that happen \emph{within} populations,
because such details are necessarily complicated, vary greatly by
application, and any attempt to encompass such a broad range of biological
processes is doomed to failure. An ``individual'' (i.e., a discrete
organism) is an abstraction that should have a well defined interpretation
in most cases, despite the endless creativity of biological reality.
This narrow focus on individuals and their groupings into populations
should ensure that the Demes standard is stable, and not subject to
continual development as more and more elaborations of basic processes
are added. We could never hope to fully encompass all the details
of every possible simulation (including mating system, ploidy,
genome sizes, recombination, mutation and gene conversion rates,
selection coefficients, and so on) and so we clearly demarcate the
role of the specification by including \emph{none} of them.
Such details can be specified by existing mechanisms used by
simulators. If they vary according to the demography (for example,
we have different recombination maps across populations), then
we can unambiguously define these by referring to the population
identifiers and times defined by the Demes model. Rather than attempting
to \emph{embed} this information into the demographic model, we
can define it externally and refer \emph{to} the demography.
